Technical Report

### 2018-03-23 fully-abstract-comp/technical-report.tex
Made T_A parameter of termination condition. Added TODO to add this to the rest of source semantics.
Added TODOs from last day in Leuven

We use the following step relation:

\begin{align*}
  \src{\Phi} & \; \src{\step \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) 
= ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
                                                    \lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta \tand \\
                                                    \lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr] \tand \\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi & \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \src{(\neg\callCond{\Phi,r_1,r_2,\aaddr} \tor}\\
                                                    \src{\eaddr < \aaddr + \calllen-1) \tand} \\
                                                    \withinBounds{\Phi(\pcreg)} \tand \\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi& \step \failed & & \totherwise
\end{align*}
\dominique{Note: The above does not allow executing code that is on the stack in the source machine, while this is in principle allowed in the target.
This is probably fine if we make sure a stack pointer can never be executable.}

\paragraph{Terminating computations}
We use the following notation to write that a configuration $\Phi$ successfully terminates:
\[
  \Phi\term[i] \defeq \Phi \nstep[i] \halted
\]
if we just know $\Phi$ terminates in some number of steps, then we write
\[
  \Phi\term \defeq \exists i \ldotp \term[i]
\]

@@

\subsection{Memory satisfaction}
\lau{14-02-2018: No matter what the memory satisfaction ends up looking like, remember to handle the revoked case.}
% {\color{DarkGreen}
Memory satisfaction for new worlds:
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
      W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W_{\var{heap}}}
    \end{array}
  \right.
\]

@@

\[
  \lrol = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
    \end{array}
\right\}
\]
\[
  \lror = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term
    \end{array}
\right\}
\]

@@

\[
  \stareg{v} = (v,H^\mathrm{sta}_{\stpair{\ms}{\ms}})
\]
\[
  H^\mathrm{sta}_{\stpair{\ms}{\ms}} = \{ \npair{\stpair{\ms}{\ms}} \}
\]

@@

\section{Full Abstraction}
\begin{definition}
  \label{def:points-to-instr}
  A configuration $(\ms,\reg)$ {\sourcecolor or $\Phi = (\ms,\reg,\_,\_)$} is said to point to non-$\scall{}{}{}$ instruction $i$ in $A$ iff
  \[
    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand \aaddr \in A
  \]
and
\[
  \decInstr{\ms(\aaddr)} = i
\]
\end{definition}

@@

\begin{definition}[Source language contextual equivalence]
  In the source language, we define that $\src{\program_1 \sconeq \program_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \term^{\ta_1}} \Leftrightarrow \src{\plug{\context}{\program_2} \term^{\ta_2}}
  \end{equation*}
  with $\src{\ta_i} = \src{\dom(\program_i.\mscode)}$
\end{definition}

>>>>

We use the following step relation:
\lau{TODO: Add $\ta$ to the source step.}
\begin{align*}
  \src{\Phi} & \; \src{\step \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) 
= ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
                                                    \lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta \tand \\
                                                    \lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr] \tand \\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi & \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \src{(\neg\callCond{\Phi,r_1,r_2,\aaddr} \tor}\\
                                                    \src{\eaddr < \aaddr + \calllen-1) \tand} \\
                                                    \withinBounds{\Phi(\pcreg)} \tand \\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi& \step \failed & & \totherwise
\end{align*}
\dominique{Note: The above does not allow executing code that is on the stack in the source machine, while this is in principle allowed in the target.
This is probably fine if we make sure a stack pointer can never be executable.}

\paragraph{Terminating computations}
We use the following notation to write that a configuration $\Phi$ successfully terminates:
\[
  \Phi\term[i] \defeq \Phi \nstep[i] \halted
\]
if we just know $\Phi$ terminates in some number of steps, then we write
\[
  \Phi\term \defeq \exists i \ldotp \term[i]
\]
We use a similar notation for source configuration termination
\[
  \Phi\sterm[i]{\ta} \defeq \Phi \nstep[i] \halted
\]
if we just know $\Phi$ terminates in some number of steps, then we write
\[
  \Phi\sterm{\ta} \defeq \exists i \ldotp \sterm[i]{\ta}
\]

@@

\subsection{Memory satisfaction}
\lau{14-02-2018: No matter what the memory satisfaction ends up looking like, remember to handle the revoked case.}
% {\color{DarkGreen}
\lau{TODO add frame to memory satisfaction and remove it from observation relation.}
Memory satisfaction for new worlds:
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
      W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W_{\var{heap}}}
    \end{array}
  \right.
\]

@@

\[
  \lrol = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta} \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
    \end{array}
\right\}
\]
\[
  \lror = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term{\ta}
    \end{array}
\right\}
\]

@@

\[
  \stareg{v} = (v,H^\mathrm{sta}_{\stpair{\ms}{\ms}})
\]
\[
  H^\mathrm{sta}_{\stpair{\ms}{\ms}} = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
    \begin{array}{l}
      \exists S : \dom(\ms) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
      \quad \forall \aaddr \in \dom(\ms) \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
    \end{array}
\right\}
\]

@@

\section{Full Abstraction}
\lau{TODOs from last day in Leuven}
\lau{TODO: components have optional main (pc/data pair)}
\lau{TODO: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
\lau{TODO: Component logical relation.}
\lau{TODO: Version of FTLR for programs.}
\lau{TODO: Component lemma for initial executable configurations.}
\lau{TODO: Executable configuration lemma.}
\lau{TODO: ... should use initial executable configuration.}
\begin{definition}
  \label{def:points-to-instr}
  A configuration $(\ms,\reg)$ {\sourcecolor or $\Phi = (\ms,\reg,\_,\_)$} is said to point to non-$\scall{}{}{}$ instruction $i$ in $A$ iff
  \[
    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand \aaddr \in A
  \]
and
\[
  \decInstr{\ms(\aaddr)} = i
\]
\end{definition}

@@

\begin{definition}[Source language contextual equivalence]
  In the source language, we define that $\src{\program_1 \sconeq \program_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \sterm{\ta[,1]}} \Leftrightarrow \src{\plug{\context}{\program_2} \sterm{\ta[,2]}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\program_i.\mscode)}$
\end{definition}
